---
title: "one_vs_group"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{one_vs_group}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval=TRUE, echo=FALSE}
library(neuroSCC)
```

# Introduction

This vignette illustrates a minimal example of a **single patient vs. group** (1 vs Group) SCC analysis with the `neuroSCC` package.

Ensure you have executed the steps from the ["Getting Started"](landing_vignette.html) vignette, or follow the conditional steps below to create required data.

---

# Data Preparation (Conditional)

If you haven't executed the previous vignette, let's conditionally recreate necessary matrices.

```{r}
cat("Creating required matrices from sample data...\n")

databaseControls <- databaseCreator(pattern = "^syntheticControl.*\\.nii\\.gz$", control = TRUE, quiet = TRUE)
databasePathological <- databaseCreator(pattern = "^syntheticPathological1\\.nii\\.gz$", control = FALSE, quiet = TRUE)

matrixControls <- matrixCreator(databaseControls, paramZ = 35, quiet = TRUE)
matrixPathological <- matrixCreator(databasePathological, paramZ = 35, quiet = TRUE)

normalizedMatrix <- meanNormalization(matrixControls)
normalizedPathological <- meanNormalization(matrixPathological)

niftiPath <- system.file("extdata", "syntheticControl1.nii.gz", package = "neuroSCC")
contours <- neuroContour(niftiPath, paramZ = 35, levels = c(0), plotResult = FALSE)

cat("Data prepared successfully.\n")
```

---

# Single Patient Data and Cloning

Prepare data for a single patient and generate synthetic Poisson clones to enable SCC computation:

```{r}
# Select the single pathological patient
SCC_AD <- normalizedPathological[1, , drop = FALSE]
SCC_CN <- normalizedMatrix

# Toy parameters for Poisson clone generation (modifiable by user)
numClones <- 2
factorLambda <- 0.1

# Generate synthetic clones
SCC_AD_clones <- generatePoissonClones(SCC_AD, numClones, factorLambda)

# Combine the patient with generated clones
SCC_AD_expanded <- rbind(SCC_AD, SCC_AD_clones)
SCC_AD_expanded <- meanNormalization(SCC_AD_expanded)
```

---

# SCC Estimation (Conditional Execution)

The SCC computation is typically computationally intensive. Here, we skip actual computation and load a precomputed result from the package (`SCCcomp.RData`):

```{r eval = requireNamespace("ImageSCC", quietly = TRUE), echo=TRUE}
result_file <- "SCCcomp.RData"
scccomp_path <- system.file("data", result_file, package = "neuroSCC")

if (file.exists(scccomp_path)) {
  cat("Precomputed SCC comparison found. Loading it...\n")
  load(scccomp_path)
} else {
  cat("Precomputed SCC data not found. Skipping due to computational complexity.\n")
  
  # Intensive SCC estimation (code provided for reference)
  SCCcomp <- ImageSCC::scc.image(
    Ya = SCC_AD_expanded,
    Yb = SCC_CN,
    Z = contours[[1]],
    d.est = 5,
    d.band = 2,
    r = 1,
    V.est.a = as.matrix(contours[[1]]),
    Tr.est.a = as.matrix(contours[[1]]),
    V.band.a = as.matrix(contours[[1]]),
    Tr.band.a = as.matrix(contours[[1]]),
    penalty = TRUE,
    lambda = 10^{seq(-6, 3, 0.5)},
    alpha.grid = c(0.10, 0.05, 0.01),
    adjust.sigma = TRUE
  )
  
  save(SCCcomp, file = result_file)
}

```

---

# Extracting Significant Points and Evaluating Metrics

Once the SCC computation results (`SCCcomp`) are available, we can extract significant points and compute relevant performance metrics:

```{r eval=TRUE}
# Extract significant points
significantPoints <- getPoints(SCCcomp)

# Load example true ROI data from the package
roi_path <- system.file("extdata", "ROIsample_Region2_18.nii.gz", package = "neuroSCC")
trueROI <- processROIs(roi_path, region = "Region2", number = "18", save = FALSE)

# Get dimensions for total coordinates
dimensions <- getDimensions(roi_path)
totalCoords <- expand.grid(x = 1:dimensions$xDim, y = 1:dimensions$yDim)

# Calculate performance metrics
metrics <- calculateMetrics(
  detectedPoints = significantPoints$positivePoints,
  truePoints = trueROI,
  totalCoords = dimensions,
  regionName = "SinglePatient_vs_Group"
)

# Display metrics
print(metrics)
```

---

# Conclusion

This vignette provides a reproducible example of conducting a **single patient vs group SCC analysis** using `neuroSCC`. Due to computational constraints, actual SCC computations are skipped and precomputed data is loaded instead. Users are encouraged to adapt parameters and explore further on their datasets.
