---
title: "two_group_comparison"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{two_group_comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval=TRUE, echo=FALSE}
library(neuroSCC)
```

# Introduction

This vignette demonstrates a minimal example of performing a **group vs. group** SCC analysis using the `neuroSCC` package.

It is recommended that you first complete the ["Getting Started"](landing_vignette.html) vignette. If you haven't, the following section will guide you through the necessary data preparation.

---

# Data Preparation (Conditional)

First, let's conditionally prepare the required data (control and pathological matrices). We use sample data included in the package.

```{r}

# Recreate necessary data from sample files
cat("Creating sample data for 2-group comparison...\n")

# Create databases for Control and Pathological groups
databaseControls <- databaseCreator(pattern = "^syntheticControl[12]\\.nii\\.gz$", control = TRUE, quiet = TRUE)
databasePathological <- databaseCreator(pattern = "^syntheticPathological[12]\\.nii\\.gz$", control = FALSE, quiet = TRUE)

# Create matrices suitable for SCC
matrixControls <- matrixCreator(databaseControls, paramZ = 35, quiet = TRUE)
matrixPathological <- matrixCreator(databasePathological, paramZ = 35, quiet = TRUE)

# Normalize matrices
normalizedControls <- meanNormalization(matrixControls)
normalizedPathological <- meanNormalization(matrixPathological)

# Extract contours from a control subject for SCC triangulation
niftiPath <- system.file("extdata", "syntheticControl1.nii.gz", package = "neuroSCC")
contours <- neuroContour(niftiPath, paramZ = 35, levels = c(0), plotResult = FALSE)

cat("Data preparation completed successfully.\n")

```

---

# SCC Estimation (Group vs Group)

The SCC estimation step is computationally intensive and thus typically skipped in vignette demonstrations. Below, we conditionally load a precomputed SCC result provided with the package.

```{r eval = requireNamespace("ImageSCC", quietly = TRUE), echo=TRUE}
if (!requireNamespace("ImageSCC", quietly = TRUE)) {
  cat("ImageSCC package is not installed.\nInstall it using: remotes::install_github('funstatpackages/ImageSCC')\n")
} else {
  result_file <- "SCCcomp.RData"
  scccomp_path <- system.file("data", result_file, package = "neuroSCC")

  if (file.exists(scccomp_path)) {
    cat("Precomputed SCC group comparison found. Loading data...\n")
    load(scccomp_path)
  } else {
    cat("Precomputed SCC data not found. Actual computation is skipped due to computational complexity.\n")

    # The following code shows how to perform the SCC computation (skipped by default)
    SCCcomp <- ImageSCC::scc.image(
      Ya = normalizedPathological,
      Yb = normalizedControls,
      Z = contours[[1]],
      d.est = 5,
      d.band = 2,
      r = 1,
      V.est.a = as.matrix(contours[[1]]),
      Tr.est.a = as.matrix(contours[[1]]),
      V.band.a = as.matrix(contours[[1]]),
      Tr.band.a = as.matrix(contours[[1]]),
      penalty = TRUE,
      lambda = 10^{seq(-6, 3, 0.5)},
      alpha.grid = c(0.10, 0.05, 0.01),
      adjust.sigma = TRUE
    )

    save(SCCcomp, file = result_file)
  }
}

```

---

# Extracting Significant Points and Evaluating Metrics

With the SCC results (`SCCcomp`) loaded, we extract the significant regions and evaluate their performance against known ground truth ROIs.

```{r eval=TRUE}
# Extract significant points from SCC results
significantPoints <- getPoints(SCCcomp)

# Load true ROI data provided within the package
roi_path <- system.file("extdata", "ROIsample_Region2_18.nii.gz", package = "neuroSCC")
trueROI <- processROIs(roi_path, region = "Region2", number = "18", save = FALSE)

# Get total coordinate dimensions
dimensions <- getDimensions(roi_path)
totalCoords <- expand.grid(x = 1:dimensions$xDim, y = 1:dimensions$yDim)

# Calculate metrics to evaluate detection performance
metrics <- calculateMetrics(
  detectedPoints = significantPoints$positivePoints,
  truePoints = trueROI,
  totalCoords = dimensions,
  regionName = "Group_vs_Group"
)

# Display calculated metrics
print(metrics)
```

---

# Conclusion

In this vignette, you've seen a concise example of conducting a **group vs group SCC analysis** using the `neuroSCC` package. We've used sample data provided by the package to illustrate the workflow clearly, focusing on practical application and evaluation of results. 

Actual SCC computations are computationally intensive and thus skipped in this vignette. Instead, we've loaded precomputed results. You can adapt this template for your own datasets, adjusting parameters and experimenting further according to your specific research needs.
